---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Emily H. K."
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r}
library(rethinking);library(ggplot2)
#Riccardo 3 C out of 6
#Make grid
dens = 20
p_grid <- seq(from = 0, to =1, length.out = dens)

#define prior
prior <- rep(1, 20)

#compute likelihood at each value in grid
likelihood <- dbinom(3, size = 6, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
plot(p_grid, posterior, type = "b")

Ricc=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
Riccplot <- ggplot(Ricc,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  ggtitle("Riccardo") + xlab("probability of knowledge")+ ylab("posterior probability")
Riccplot

#Implement quadractic aprox.
#If we assume that distrubution is gaussian, all we need is the mean and SD. From there we cab shape a distrubution. 
Ricc.qa <- map(
  alist(
    w ~ dbinom(6,p),
    p ~ dunif(0,1)
  ) ,
  data=list(w=3))

precis(Ricc.qa)

#See how much of the porsterior distrub that falls above 0.5.
sum(posterior[p_grid < 0.5])
#There is a 50% probability that Riccardo would know more than chance

```


2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.
```{r}
#- Kristian: 2 correct answers out of 2 questions (then he gets bored)
#- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
#- Mikkel: 66 correct answers out of 132 questions

#Kristian
#Grid
dens = 20
p_grid <- seq(from = 0, to =1, length.out = dens)

#define prior
prior <- rep(1, 20)
#compute likelihood at each value in grid
likelihood <- dbinom(2, size = 2, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Kris=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
Krisplot <- ggplot(Kris,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Kristian") + xlab("probability of knowledge")+ ylab("posterior probability")

#Josh
#P grid for josh

p_gridJ <- seq(from = 0, to =1, length.out = 20)

#compute likelihood at each value in grid
likelihood <- dbinom(160, size = 198, prob = p_gridJ)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Josh=data.frame(grid=p_gridJ,posterior=posterior,prior=prior,likelihood=likelihood)
Joshplot<- ggplot(Josh,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Joshua")+ xlab("probability of knowledge")+ ylab("posterior probability")


#Mikkel
#grid
p_gridM <- seq(from = 0, to =1, length.out = 20)
#compute likelihood at each value in grid
likelihood <- dbinom(66, size = 132, prob = p_gridM)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Mikkel=data.frame(grid=p_gridM,posterior=posterior,prior=prior,likelihood=likelihood)
Mikkelplot <- ggplot(Mikkel,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Mikkel")+  xlab("probability of knowledge")+ ylab("posterior probability") 

#Look at riccardo and mikkel plots
gridExtra::grid.arrange(Riccplot, Mikkelplot)
gridExtra::grid.arrange(Krisplot,Joshplot, Riccplot, Mikkelplot)
Mikk.qa <- map(
  alist(
    w ~ dbinom(132,p),
    p ~ dunif(0,1)
  ) ,
  data=list(w=66))

precis(Mikk.qa)
precis(Ricc.qa)
sum(posterior[p_gridM > 0.6])

```
Although Riccardo and Mikkel have the same beta, we can see that there is a lot of uncertainty for Riccardos knowledge above chance. If we look at Mikkels we see that there is much less variance, witha  SD of 0.2.

3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.
```{r}
#Change prior of teachers
prior = dnorm(p_grid, mean = 0.8, sd = 0.2)

#Kristian
#Grid
dens = 20
p_grid <- seq(from = 0, to =1, length.out = dens)

#compute likelihood at each value in grid
likelihood <- dbinom(2, size = 2, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Kris2=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
Krisplot2 <- ggplot(Kris2,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  ggtitle("Kristian") + xlab("probability of knowledge")+ ylab("posterior probability")
Krisplot2

#Josh
#P grid for josh

p_gridJ <- seq(from = 0, to =1, length.out = 20)

#compute likelihood at each value in grid
likelihood <- dbinom(160, size = 198, prob = p_gridJ)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Josh2=data.frame(grid=p_gridJ,posterior=posterior,prior=prior,likelihood=likelihood)
Joshplot2<- ggplot(Josh2,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Joshua") + xlab("probability of knowledge")+ ylab("posterior probability")
Joshplot2

#Mikkel
#grid

p_gridM <- seq(from = 0, to =1, length.out = 20)
#compute likelihood at each value in grid
likelihood <- dbinom(66, size = 132, prob = p_gridM)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Mikkel2=data.frame(grid=p_gridM,posterior=posterior,prior=prior,likelihood=likelihood)
Mikkelplot2 <- ggplot(Mikkel2,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  ggtitle("Mikkel") + xlab("probability of knowledge")+ ylab("posterior probability") 
Mikkelplot2

#Riccardo
#grid
p_gridR <- seq(from = 0, to =1, length.out = 20)

#likelihood
likelihood <- dbinom(3, size = 6, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Ricc2=data.frame(grid=p_gridR,posterior=posterior,prior=prior,likelihood=likelihood)
Riccplot2 <- ggplot(Ricc2,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Riccardo")+ xlab("probability of knowledge")+ ylab("posterior probability")
Riccplot2

sum(posterior[p_gridR < 0.5])



gridExtra::grid.arrange(Krisplot2, Joshplot2, Riccplot2, Mikkelplot2)

gridExtra::grid.arrange(Mikkelplot2, Riccplot2)
```

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?
```{r}

#Uniform prior

# - Kristian: 2*100 correct answers out of 2*100 questions (then he gets bored)
# - Josh: 160*100 correct answers out of 198*100 questions (Josh never gets bored)
# - Riccardo: 3*100 correct answers out of 6*100 questions
# - Mikkel: 66*100 correct answers out of 132*100 questions

#Kristian
#Grid
dens = 200
p_grid <- seq(from = 0, to =1, length.out = dens)

#define prior
prior <- rep(1, 200)
#compute likelihood at each value in grid
likelihood <- dbinom(200, size = 200, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Standarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Kris3=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
Krisplot3 <- ggplot(Kris3,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Kristian") + xlab("probability of knowledge")+ ylab("posterior probability")

#Josh
#P grid for josh

p_gridJ <- seq(from = 0, to =1, length.out = dens)

#compute likelihood at each value in grid
likelihood <- dbinom(16000, size = 19800, prob = p_gridJ)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Josh3=data.frame(grid=p_gridJ,posterior=posterior,prior=prior,likelihood=likelihood)
Joshplot3<- ggplot(Josh3,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Joshua")+ xlab("probability of knowledge")+ ylab("posterior probability")


#Riccardo
#grid
p_gridR <- seq(from = 0, to =1, length.out = dens)
#compute likelihood at each value in grid
likelihood <- dbinom(300, size = 600, prob = p_gridM)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Riccardo3=data.frame(grid=p_gridR,posterior=posterior,prior=prior,likelihood=likelihood)
Riccardoplot3 <- ggplot(Riccardo3,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Riccardo")+  xlab("probability of knowledge")+ ylab("posterior probability") 


#Mikkel
#grid
p_gridM <- seq(from = 0, to =1, length.out = dens)
#compute likelihood at each value in grid
likelihood <- dbinom(6600, size = 13200, prob = p_gridM)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Mikkel3=data.frame(grid=p_gridM,posterior=posterior,prior=prior,likelihood=likelihood)
Mikkelplot3 <- ggplot(Mikkel3,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Mikkel")+  xlab("probability of knowledge")+ ylab("posterior probability") 


gridExtra::grid.arrange(Krisplot3,Joshplot3, Riccardoplot3, Mikkelplot3)

#_________________________________________________________________________________________
#Mean prior, mean = 0.8, SD = 0.2

prior = dnorm(p_grid, mean = 0.8, sd = 0.2)


#Kristian
#Grid
dens = 200
p_grid <- seq(from = 0, to =1, length.out = dens)

#compute likelihood at each value in grid
likelihood <- dbinom(200, size = 200, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Kris3a=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
Krisplot3a <- ggplot(Kris3a,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Kristian") + xlab("probability of knowledge")+ ylab("posterior probability")

#Josh
#P grid for josh

p_gridJ <- seq(from = 0, to =1, length.out = dens)

#compute likelihood at each value in grid
likelihood <- dbinom(16000, size = 19800, prob = p_gridJ)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Josh3a=data.frame(grid=p_gridJ,posterior=posterior,prior=prior,likelihood=likelihood)
Joshplot3a<- ggplot(Josh3a,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Joshua")+ xlab("probability of knowledge")+ ylab("posterior probability")


#Riccardo
#grid
p_gridR <- seq(from = 0, to =1, length.out = dens)
#compute likelihood at each value in grid
likelihood <- dbinom(300, size = 600, prob = p_gridM)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Riccardo3a=data.frame(grid=p_gridR,posterior=posterior,prior=prior,likelihood=likelihood)
Riccardoplot3a <- ggplot(Riccardo3a,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Riccardo")+  xlab("probability of knowledge")+ ylab("posterior probability") 


#Mikkel
#grid
p_gridM <- seq(from = 0, to =1, length.out = dens)
#compute likelihood at each value in grid
likelihood <- dbinom(6600, size = 13200, prob = p_gridM)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Mikkel3a=data.frame(grid=p_gridM,posterior=posterior,prior=prior,likelihood=likelihood)
Mikkelplot3a <- ggplot(Mikkel3a,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Mikkel")+  xlab("probability of knowledge")+ ylab("posterior probability") 


gridExtra::grid.arrange(Krisplot3a,Joshplot3a, Riccardoplot3a, Mikkelplot3a)

#___________________________________________________________________________________
#Change prior of teachers
prior = dnorm(p_grid, mean = 0.8, sd = 0.2)

#Kristian
#Grid
dens = 200
p_grid <- seq(from = 0, to =1, length.out = dens)

#compute likelihood at each value in grid
likelihood <- dbinom(200, size = 200, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Kris2=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
Krisplot2 <- ggplot(Kris2,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  ggtitle("Kristian") + xlab("probability of knowledge")+ ylab("posterior probability")
Krisplot2

#Josh
#P grid for josh

p_gridJ <- seq(from = 0, to =1, length.out = dens)

#compute likelihood at each value in grid
likelihood <- dbinom(16000, size = 19800, prob = p_gridJ)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Josh2=data.frame(grid=p_gridJ,posterior=posterior,prior=prior,likelihood=likelihood)
Joshplot2<- ggplot(Josh2,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Joshua") + xlab("probability of knowledge")+ ylab("posterior probability")
Joshplot2

#Mikkel
#grid

p_gridM <- seq(from = 0, to =1, length.out = dens)
#compute likelihood at each value in grid
likelihood <- dbinom(6600, size = 13200, prob = p_gridM)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Mikkel2=data.frame(grid=p_gridM,posterior=posterior,prior=prior,likelihood=likelihood)
Mikkelplot2 <- ggplot(Mikkel2,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+  ggtitle("Mikkel") + xlab("probability of knowledge")+ ylab("posterior probability") 
Mikkelplot2

#Riccardo
#grid
p_gridR <- seq(from = 0, to =1, length.out = dens)

#likelihood
likelihood <- dbinom(300, size = 600, prob = p_grid)

#compute product of likelihood and prior
unstd.posterior <- likelihood*prior
#Starndarize the posterior so it sums 1
posterior <- unstd.posterior/sum(unstd.posterior)

#Plot
Ricc2=data.frame(grid=p_gridR,posterior=posterior,prior=prior,likelihood=likelihood)
Riccplot2 <- ggplot(Ricc2,aes(grid,posterior))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='red')+ ggtitle("Riccardo")+ xlab("probability of knowledge")+ ylab("posterior probability")
Riccplot2

gridExtra::grid.arrange(Krisplot2, Joshplot2, Mikkelplot2, Riccplot2)




```

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?
```{r}
#I would use a sceptical prior, below 0.5. There would be a need of a lot of evidence for going above .50
```

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?
```{r}

```

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



